\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-07-03 forwards-and-backwards.tex
.
peter edits

As one takes a journey, long or short, the landscape changes:
forests thicken or thin or change their composition,
springtime grasslands host intergrading mosaics of different types of flowers.
The aim of this paper is to understand the implications of a broad class of mechanistic spatial models
that might describe how such populations live, die, and reproduce:
How does population density change across space and time?
And, how might we learn about the underlying dynamics from genealogical or genetic data?
We work with individual-based models of a single species in continuous space,
where birth, death and establishment may all depend on spatial location
as well as on local population density, allowing for stable populations through density-dependent feedback.

There is now a huge literature devoted to spatial branching processes,
including branching random walk, 
branching Brownian motion, and the Dawson-Watanabe superprocesses.  
In addition to possessing a rich and beautiful mathematical structure, 
the superprocesses are remarkable for demonstrating a certain `universality', 
arising, as they do, as scaling limits of a vast array of different 
spatial population models \cite[e.g.,][]{chetwynd-diggle/etheridge:2018,
cox/perkins:2005,
cox/durrett/perkins:1999,
cox/durrett/perkins:2000,
holmes:2008,
vanderhofstad/sakai:2010,
vanderhofstad/slade:2003,
vanderhofstad/holmes/perkins:2017}.

@@

The porous medium equation (PME)
is a degenerate nonlinear diffusion equation that
has been the subject of intensive study. We refer to~\cite{vazquez:2007} for 
a comprehensive reference. The equation has also been studied with various
forms of noise, see~\cite{barbu/daprato/roeckner:2016}, although not the 
form of noise that will arise from the population dynamics below. 
Deriving nonlinear diffusion equations as limits of weakly interacting 
diffusions goes back at least to~\cite{mckean:1967}, and this 
programme has been
carried out for the PME and its generalisations. 
The degeneracy of the PME
results in some extra analytic
challenges,~\cite{jourdain:2000}.
Recent work in this 
direction has been spurred on by applications
in mathematical finance, %where the porous medium arises 
in particular, in the study of weak limits of systems
of diffusions interacting through their 
rank,~\cite{dembo/shkolnikov/varadhan/zeitouni:2016, jourdain/reygner:2013}
and references therein.
Our approach is different.
We start from individual
based population models, thus providing a microscopic mechanism that leads to 
nonlinear diffusion.
Other microscopic mechanisms have been proposed in the special case of the
PME, see 
e.g.~\cite{feng/iscoe/seppalainen:1997, oelschlaeger:1990}; ours
is specifically motivated by biological considerations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Notes to ourselves}

\plr{For recording/discussing notation decisions.}

$X$ is the original, unscaled process,
and $\eta(t) = X(t\theta)/N$.

When the process has a density, it is $\varphi(t, x) = d \eta_t(dx)/dx$.

$$\lim \frac{\theta}{N} =: \alpha$$

Were not using $\lambda$, but only $N$.

@@

Birth, establishment, and death can depend on the location of the individual
and the local population density.
Since we would like the population density to scale with $N$,
these are functions of $X/N$, i.e.,
the counting measure with mass $1/N$ placed at the location of each individual.
To measure local population density, we convolve the population measure with
an ``interaction kernel'' -- which, for simplicity, we take to be Gaussian,
which we write as
$$
    p_{\epsilon^2}(x) := \frac{1}{(2 \pi \epsilon^2)^{d/2}} e^{-\|x\|^2 / 2 \epsilon^2} ,
$$
for an ``interaction distance'' $\epsilon$.
First consider birth rates, defined by
a nonnegative function $\gamma(x, m) : \IR^d \times \IR_{\ge 0} \to \IR_{\ge 0}$
of location $x$ and local population density $m$,
and an interaction distance $\epsilon_\gamma > 0$.
Local population density enters through the convolution of $X$ with $p_{\epsilon_\gamma^2}$,
which we write as $\smooth{\gamma} X$.
Then, the birth rate of an individual at location $x$ when the state of the population is $X$
is $\gamma(x, \smooth{\gamma} X(x) / N)$.
Similarly, the establishment probability of an offspring at location $x$ is
is $r(x, \smooth{r} X(x) / N)$,
where $r(x, m) : \IR^d \times \IR_{\ge 0} \to [0, 1]$
and again $\smooth{r} X$ is the convolution of $p_{\epsilon_r^2}$ with $X$.
Rather than defining a function $\mu(x, m)$ for the death rate,
it turns out to be more convenient to parameterize in terms of the net reproductive rate,
which we denote by $F(x, m)$,
scaled by $1/\theta$
so that the population density changes over a time scale of $\theta$.
(This corresponds to a ``near-critical branching'' assumption.)
Informally, we would like that
$$
    F = \theta (r \gamma - \mu) ,
$$
and to arrange this we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where again $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR_{\ge 0}$
and $\kernel_F = p_{\epsilon_F^2}$.

In summary, each of the three demographic parameters depend on local density measured
with a Gaussian kernel with possibly different scales:
for birth over a scale of $\epsilon_\gamma$,
for establishment over a scale of $\epsilon_r$,
and for net reproductive rate over a scale of $\epsilon_F$.
Note that death rates depend in principle on population densities at three different scales,
so that we could write $\mu = \mu(x, \smooth{\gamma} X(x) / N, \smooth{r} X(x) / N, \smooth{F} X(x) / N)$.
It might seem like
we have introduced unnecessary complexity
with these three different spatial scales on which interactions occur.
However,
it will turn out to be useful in the proofs to have $\epsilon_r < \epsilon_\gamma$,
although in the limit they can be taken equal to each other
(and $\epsilon_F$ is unconstrained),
so informally we think of local density as being determined on a scale of $\epsilon$.

@@

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
\end{definition}

Expressions like $\gamma(x, \smooth{\gamma} \eta(x))$ will appear often in the paper.
To make formula more readable, we overload notation to define
$$
    \gamma(x, \eta) := \gamma(x, \smooth{\gamma} \eta(x)) ,
$$
and similarly write $\mu(x, \eta)$  for $\mu(x, \smooth{\mu} \eta(x))$ and $r(x, \eta)$ for $r(x, \smooth{r} \eta(x))$.
When convenient, we may also suppress the arguments completely,
writing simply $\gamma$, $\mu$, and $r$ for these quantities.

For each $N > 0$ and $\theta \ge 1/(\epsilon_\gamma - \epsilon_r)$,
the process $\left(\eta^N(t)\right)_{t \ge 0}$
is the continuous-time measure-valued Markov process described above
with birth rates, establishment probabilities, and death rates
determined as follows.
When the state of the process is $\eta \in \measures$,
offspring of mass $1/N$ are produced at location $x$
with rate $\theta \gamma(x, \eta) N \eta_s(dx)$
and adults die at location $x$ with rate $\theta \mu(x, \eta) N \eta_s(dx)$,
(i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
Each offspring disperses to a location offset from the parent's location
by an independent Gaussian with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$;
an offspring at $x$ establishes instantaneously
with probability $r(x, \eta)$, or else dies.

We have just effectively defined the process, but to be completely clear,
we will now give a formal definition as a solution to a martingale problem.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
We define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{2}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eq: Pre-limit Martingale Characterisation}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
\begin{equation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } N \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) .
    $$
\end{definition}

\paragraph{Remark:}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(x, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(x, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
should be biased towards regions of higher establishment probability.
For comparison to future results,
it is interesting to write down the limiting generator:
\begin{align*}
    \lim_{N \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(\cdot) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align*}
In the simplest case of unbiased, isotropic dispersal,
$\meanq = 0$ and $\covq = I$, so $\DG = \Delta$,
and this is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
However, we will see that following a lineage back through the population
has a different behavior yet again.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, $N \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)

    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f:\mathbb{R}^d \rightarrow \mathbb{R}$,
    \begin{align} \label{eqn:mgale_problem}
        \begin{split}
        M_t
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \smooth{\gamma} \eta_{s}(x))
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \smooth{r} \eta_{s}(\cdot))
                \right)(x)
                \\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \smooth{F} \eta_{s}(x)),
                \eta_{s}(dx)
            \big\rangle ds,
        \end{split}
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x,  \eta_{s} \right)
            r\left(x, \eta_{s} \right)+\mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

Theorem \ref{thm:nonlocal_convergence} provided convergence along subsequences;
i.e., for any sequence of values of $\theta$
there is a subsequence that converges.
However, if the limiting process is unique then the convergence is unqualified.
In the deterministic case, we have convergence to a nonlocal PDE.

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}
    has a unique solution,
    then $(\eta^N(t))_{t \ge 0}$ converges to that solution
    as $N \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

Informally, equation \eqref{eqn:nonlocal_pde} is the PDE
\begin{align} \label{eqn:pde}
    \dot \varphi = r \DG^* \left( \gamma \varphi \right) + \varphi F ,
\end{align}
where $\DG$ is an elliptic second order differential operator,
and the coefficients $r$, $\gamma$, and $F$
at $x$ depend nonlocally on $\varphi$
(i.e., through a smoothed version of $\varphi$).
To see why it is $\DG^*$ and not $\DG$ that appears in this equation,
consider the case where dispersal has a mean drift to the right,
so that the drift term in $\DG$ is $d/dx$.
Then, the population density at a location with positive slope
will tend to \emph{decrease}, i.e., proportional to $-d/dx$,
because the density at a location is determined by the offspring that disperse \emph{to}
that location, not away from it.
Furthermore, if dispersal tends to move offspring away from a region
(i.e., if divergence $\grad \cdot \meanq > 0$)
this will tend to decrease the local density.

Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
they depend only on the region nearby to $x$,
and so we would expect solutions of the nonlocal PDE to be close to the corresponding local PDE
as the interaction distances $\epsilon_r$, $\epsilon_\gamma$ and $\epsilon_F$
go to zero:
for instance, if $\eta_t$ has density $\varphi_t$
then $\gamma(x, \smooth{\gamma} \eta_t(x)) \to \gamma(x, \varphi_t(x))$ as $\epsilon_\gamma \to 0$.
The following propositions give two concrete situations in which that is true.

>>>>

\comment{TO CITE: \citet{ghosh2022emergent}}

As one takes a journey, long or short, the landscape changes:
forests thicken or thin or change their composition,
springtime grasslands host intergrading mosaics of different types of flowers.
The aim of this paper is to understand the implications of a broad class of mechanistic spatial models
that might describe how such spatially heterogeneous populations live, die, and reproduce:
How does population density change across space and time?
How might we learn about the underlying dynamics from genealogical or genetic data?
And,  how does genetic ancestry spread across geography,
when looking back through time in these populations?
We work with individual-based models of a single species in continuous space,
where birth, death and establishment may all depend on spatial location
as well as on local population density, allowing for stable populations through density-dependent feedback.

Reproduction of individuals naturally leads to a branching process.
There is now a huge literature devoted to spatial branching processes,
including branching random walk, 
branching Brownian motion, and the Dawson-Watanabe superprocesses.  
In addition to possessing a rich and beautiful mathematical structure, 
the superprocesses are remarkable for demonstrating a certain `universality', 
arising, as they do, as scaling limits of a vast array of different 
spatial population models \cite[e.g.,][]{chetwynd-diggle/etheridge:2018,
cox/perkins:2005,
cox/durrett/perkins:1999,
cox/durrett/perkins:2000,
holmes:2008,
vanderhofstad/sakai:2010,
vanderhofstad/slade:2003,
vanderhofstad/holmes/perkins:2017}.

@@

The porous medium equation (PME)
is a degenerate nonlinear diffusion equation that
has been the subject of intensive study. We refer to~\cite{vazquez:2007} for 
a comprehensive reference. The equation has also been studied with various
forms of noise, see~\cite{barbu/daprato/roeckner:2016}, although not the 
form of noise that will arise from the population dynamics below. 
Deriving nonlinear diffusion equations as limits of weakly interacting 
diffusions goes back at least to~\cite{mckean:1967}, and this 
programme has been
carried out for the PME and its generalisations. 
The degeneracy of the PME
results in some extra analytic
challenges,~\cite{jourdain:2000}.
% Recent work in this 
% direction has been spurred on by applications
% in mathematical finance, %where the porous medium arises 
% in particular, in the study of weak limits of systems
% of diffusions interacting through their 
% rank,~\cite{dembo/shkolnikov/varadhan/zeitouni:2016, jourdain/reygner:2013}
% and references therein.
Our approach is different.
We start from individual
based population models, thus providing a microscopic mechanism that leads to 
nonlinear diffusion.
Other microscopic mechanisms have been proposed in the special case of the
PME, see 
e.g.~\cite{feng/iscoe/seppalainen:1997, oelschlaeger:1990}; ours
is specifically motivated by biological considerations.

@@

@@

Birth, establishment, and death can depend on the location of the individual
and the local population density.
Since we would like the population density to scale with $N$,
these are functions of $X/N$, i.e.,
the counting measure with mass $1/N$ placed at the location of each individual.
To measure local population density, we convolve the population measure with
an ``interaction kernel'' -- which, for simplicity, we take to be Gaussian,
and we write as
$$
    p_{\epsilon^2}(x) := \frac{1}{(2 \pi \epsilon^2)^{d/2}} e^{-\|x\|^2 / 2 \epsilon^2} ,
$$
for an ``interaction distance'' $\epsilon$.
First consider birth rates, defined by
a nonnegative function $\gamma(x, m) : \IR^d \times \IR_{\ge 0} \to \IR_{\ge 0}$
of location $x$ and local population density $m$,
and an interaction distance $\epsilon_\gamma > 0$.
Local population density enters through the convolution of $X$ with $p_{\epsilon_\gamma^2}$,
which we write as $\smooth{\gamma} X$.
Then, the birth rate of an individual at location $x$ when the state of the population is $X$
is $\gamma(x, \smooth{\gamma} X(x) / N)$.
Similarly, the establishment probability of an offspring at location $x$ is
is $r(x, \smooth{r} X(x) / N)$,
where $r(x, m) : \IR^d \times \IR_{\ge 0} \to [0, 1]$
and again $\smooth{r} X$ is the convolution of $p_{\epsilon_r^2}$ with $X$.
Rather than defining a function $\mu(x, m)$ for the death rate,
it turns out to be more convenient to parameterize in terms of the net reproductive rate,
which we denote by $F(x, m)$,
scaled by $1/\theta$
so that the population density changes over a time scale of $\theta$.
(This corresponds to a ``near-critical branching'' assumption.)
Informally, we would like that
$$
    F = \theta (r \gamma - \mu) ,
$$
and to arrange this we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where again $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR_{\ge 0}$
and $\kernel_F = p_{\epsilon_F^2}$.

In summary, each of the three demographic parameters depend on local density measured
with a Gaussian kernel with possibly different scales:
for birth over a scale of $\epsilon_\gamma$,
for establishment over a scale of $\epsilon_r$,
and for net reproductive rate over a scale of $\epsilon_F$.
Note that death rate depends in principle on population densities at three different scales,
so that we could write $\mu = \mu(x, \smooth{\gamma} X(x) / N, \smooth{r} X(x) / N, \smooth{F} X(x) / N)$.
It might seem like
we have introduced unnecessary complexity
with these three different spatial scales on which interactions occur.
However,
it will turn out to be useful in the proofs to have $\epsilon_r < \epsilon_\gamma$,
although in the limit they can be taken equal to each other
(and $\epsilon_F$ is unconstrained),
so informally we think of local density as being determined on a scale of $\epsilon$.

@@

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix positive values $\theta$, $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
Although the models will make sense without this requirement,
for the purposes of convergence we will also require that $\theta > 1/(\epsilon_\gamma - \epsilon_r)$.
\end{definition}

Expressions like $\gamma(x, \smooth{\gamma} \eta(x))$ will appear often in the paper.
To make formula more readable, we overload notation to define
$$
    \gamma(x, \eta) := \gamma(x, \smooth{\gamma} \eta(x)) ,
$$
and similarly write $r(x, \eta)$ for $r(x, \smooth{r} \eta(x))$
and $\mu(x, \eta)$ for the expression of equation \eqref{eqn:mu_defn}.
When convenient, we may also suppress the arguments completely,
writing simply $\gamma$, $\mu$, and $r$ for these quantities.

For each value of $N$ and $\theta$,
the process $\left(\eta^N_t \right)_{t \ge 0}$
is the continuous-time measure-valued Markov process described above
with birth rates, establishment probabilities, and death rates
determined as follows.
When the state of the process is $\eta \in \measures$,
offspring of mass $1/N$ are produced at location $x$
with rate $\theta \gamma(x, \eta) N \eta(dx)$
and adults die at location $x$ with rate $\theta \mu(x, \eta) N \eta(dx)$,
(i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
Each offspring disperses to a location offset from the parent's location
by an independent Gaussian with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$;
an offspring at $x$ establishes instantaneously
with probability $r(x, \eta)$, or else dies.

We have just effectively defined the process, but for future use,
we will now give a formal definition as a solution to a martingale problem.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
For each value of $N$ and $\theta = \theta(N)$,
we define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{\infty}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eq: Pre-limit Martingale Characterisation}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
\begin{equation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align*}
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
        &=
        \text{TODO} .
    \end{align*}
\end{definition}

\paragraph{Remark:}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison to future results,
it is interesting to write down the limiting generator:
\begin{align*}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align*}
In the simplest case of unbiased, isotropic dispersal,
$\meanq = 0$ and $\covq = I$, so $\DG = \Delta$,
and this is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following a lineage of successful individuals back through time
would be described by this generator in the limit.
However, we will see that following a lineage back through the population
has a different behavior yet again.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, also $\theta \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)
    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f \in C^\infty_0(\IR^d)$,
    \begin{align} \label{eqn:mgale_problem}
        \begin{split}
        M_t
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                \\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds,
        \end{split}
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right) + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

Theorem \ref{thm:nonlocal_convergence} provided convergence along subsequences;
i.e., for any sequence of values of $N$
there is a subsequence along which $\eta^N$ converges.
However, if the limiting process is unique then the convergence is unqualified.
In the deterministic case, we have convergence to a nonlocal PDE.

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges to that solution
    as $N \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

Informally, equation \eqref{eqn:nonlocal_pde} is the PDE of Definition~\ref{defn:weak_solutions},
where $\DG$ is an elliptic second order differential operator,
and the coefficients $r$, $\gamma$, and $F$
at $x$ depend nonlocally on $\varphi$
(i.e., through a smoothed version of $\varphi$).
To see why it is $\DG^*$ and not $\DG$ that appears in this equation,
consider the case where dispersal has a mean drift to the right,
so that the drift term in $\DG$ is $d/dx$.
Then, the population density at a location with positive slope
will tend to \emph{decrease}, i.e., proportional to $-d/dx$,
because the density at a location is determined by the offspring that disperse \emph{to}
that location, not away from it.
Furthermore, if dispersal tends to move offspring away from a region
(i.e., if divergence $\grad \cdot \meanq > 0$)
this will tend to decrease the local density.

Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
they depend only on the region nearby to $x$,
and so we would expect solutions of the nonlocal PDE to be close to the corresponding local PDE
as the interaction distances $\epsilon_r$, $\epsilon_\gamma$ and $\epsilon_F$
go to zero:
for instance, if $\eta_t$ has density $\varphi_t$
then $\gamma(x, \smooth{\gamma} \eta_t(x)) \to \gamma(x, \varphi_t(x))$ as $\epsilon_\gamma \to 0$
(recalling that $\smooth{\gamma}{}$ denotes convolution with the heat kernel at time $\epsilon_\gamma^2$).
The following propositions give two concrete situations in which that is true.
